---
title: "Assignment04"
format: html
author: J Lei
---

The dataset used in this project is from Kaggle by a user named NIDULA ELGIRIYEWITHANA, the dataset is named [Billionaires Statistics Dataset](https://www.kaggle.com/datasets/nelgiriyewithana/billionaires-statistics-dataset) . This data records the statues of billionaires in the world and can help us to explore the topic of wealth inequality and income inequality.

# Setup

### 1) Hide unnecessary print

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### 2) Set data

```{r}
library(tidyverse)
data <- read_csv("data/Billionaires Statistics Dataset.csv")
  
  
```

# Graphs

### 1) Graph 1

```{r}
data %>%
  group_by(country) %>%
  summarise(billionaire_count = n()) %>%
  arrange(desc(billionaire_count)) %>%
  slice_head(n = 5) %>%
  ggplot(aes(x = reorder(country, -billionaire_count), y = billionaire_count, fill = country)) +
  geom_col() +
  labs(title = "Top 5 Countries with the Most Billionaires",
       x = "Country",
       y = "Number of Billionaires",
       caption = "source:Billionaires Statistics Dataset") +
  theme_minimal()

```

The first graph show the 5 countries with most billionaires are United States, China, India, Germany, and United Kingdom. This raises immediate concerns on income distribution as China and India are developing countries with far less GDP per capita comparing to the other three countries.

### 

### 2) Graph 2

```{r}
top_billionaires <- data %>%
  arrange(desc(finalWorth)) %>%
  head(20)

  top_billionaires %>%
ggplot(aes(x = country, y = finalWorth, color = category)) +
  geom_point(size = 2, alpha = 0.7) +            
  geom_text(aes(label = personName), size = 2) + 
  labs(title = "Billionaires by Country and Sector",
       x = "Country",
       y = "Final Worth",
       color = "Sector",
       caption = "source:Billionaires Statistics Dataset"
       ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  

```

The Second graph shows the top 20 billionaires' country, sector, and final worth. This graph shows that Fashion & Retail and technology are the two sectors for top billionaires across all countries concentrated in. It might suggest either those sectors are earning abnormal profit by abusing monopoly power.
