---
title: "Assignment04"
format: html
author: J Lei
---

The dataset used in this project is from Kaggle by a user named NIDULA ELGIRIYEWITHANA, the dataset is named [Billionaires Statistics Dataset](https://www.kaggle.com/datasets/nelgiriyewithana/billionaires-statistics-dataset) . This data records the statues of billionaires in the world and can help us to explore the topic of wealth inequality and income inequality.

# Setup

### 1) Hide unnecessary print

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### 2) Set data

```{r}
library(tidyverse)
data <- read_csv("data/Billionaires Statistics Dataset.csv")
  
  
```

# Graphs

### 1) Graph 1

```{r}
data %>%
  group_by(country) %>%
  summarise(billionaire_count = n()) %>%
  arrange(desc(billionaire_count)) %>%
  slice_head(n = 5) %>%
  ggplot(aes(x = reorder(country, -billionaire_count), y = billionaire_count, fill = country)) +
  geom_col() +
  labs(title = "Top 5 Countries with the Most Billionaires",
       x = "Country",
       y = "Number of Billionaires",
       caption = "source:Billionaires Statistics Dataset") +
  theme_minimal()

```

The first graph show the 5 countries with most billionaires are United States, China, India, Germany, and United Kingdom. This raises immediate concerns on income distribution as China and India are developing countries with far less GDP per capita comparing to the other three countries.

### 

### 2) Graph 2

```{r}
top_billionaires <- data %>%
  arrange(desc(finalWorth)) %>%
  head(20)

  top_billionaires %>%
ggplot(aes(x = country, y = finalWorth, color = category, size = finalWorth, alpha = 0.7)) +
  geom_point() +            
  geom_text(aes(label = personName), size = 2) + 
  labs(title = "Billionaires by Country and Sector",
       x = "Country",
       y = "Final Worth(million dollars)",
       color = "Sector",
       size = "Final Worth(million dollars)",
       caption = "source:Billionaires Statistics Dataset"
       ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

  

```

The Second graph shows the top 20 billionaires' country, sector, and final worth. This graph shows that Fashion & Retail and technology are the two sectors for top billionaires across all countries concentrated in. It might suggest either those sectors are earning abnormal profit by abusing monopoly power.

### Graph 3 

```{r message=FALSE}

library(ggthemes)

data <- data %>%
  mutate(gdp_numeric = as.numeric(gsub("[$,]", "", gdp_country))) %>%
  mutate(gdp_percap = gdp_numeric / population_country)


summary_data <- data %>%
  group_by(gdp_percap) %>%
  summarize(num_billionaires = n()) %>%
  drop_na(gdp_percap)

ggplot(summary_data, aes(x = gdp_percap, y = num_billionaires)) +
  geom_line() +
  geom_smooth (method = "lm", se = TRUE) +
  labs(title = "Number of Billionaires vs. GDP per Capita(dollars)", 
       x = "GDP per Capita(dollars)", 
       y = "Number of Billionaires",
       caption = "source:Billionaires Statistics Dataset"
       )+
  scale_x_continuous(limits = c(0,185000), 
                     breaks = seq(from = 0, to = 200000, by = 20000))+
  theme_economist()
```

The third graph is more of a in-depth analysis on the problem discovred from graph. Graph 3 is trying to see if there is correlation between income and number of billionaires. If they are highly related, number of billionaires might be less a concern and it might increase with incrase in national average income. However, although the graph shows that there might be a linear correlation, there are numbers of abnormal value.(GDP per capita is low while numbers of billionaires are extremely high)

### Graph 4 

()

```{r}
data %>%
 drop_na(total_tax_rate_country) %>%
  filter(total_tax_rate_country < 100) %>%
  ggplot(aes(x = total_tax_rate_country)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) + 
  labs(
    title = "Distribution of Tax Rates among Billionaires",
    x = "Tax Rate",
    y = "Number of Billionaires",
    caption = "source: Billionaires Statistics Dataset"
  ) +
   scale_x_continuous(limits = c(0,75), 
                     breaks = seq(from = 0, to = 75, by = 5))+
  theme_minimal()
```

Graph 4 shows how tax rate might impact the creation of billionaires. On the data set, Argentina has a total tax rate of 110 which intuitively looks suspicious, therefore it has been excluded. From this graph, it seems that there is no pattern showing high tax rate or low tax rate would make a difference on number of billionaires. It might provide some justifications for higher tax rate on the rich. However, it is a very premature hypothesis which needs to be proven with much deeper analysis.
